---
title: "E18 mLN pLN EYFP+"
author: "A.DeMartin"
date: "2025-05-23"
output: 
  html_document:
    keep_md: true
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
options(width = 100)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

### reprocess
#### load packages
```{r load packages, warning=FALSE, include=FALSE}
library(ExploreSCdataSeurat3)
library(runSeurat3)
library(Seurat)
library(ggpubr)
library(pheatmap)
library(SingleCellExperiment)
library(dplyr)
library(tidyverse)
library(viridis)
library(muscat)
library(circlize)
library(destiny)
library(scater)
library(metap)
library(multtest)
library(clusterProfiler)
library(org.Mm.eg.db)
library(msigdbr)
library(enrichplot)
library(DOSE)
library(grid)
library(gridExtra)
library(ggupset)
library(VennDiagram)
library(NCmisc)
library(slingshot)
library(here)
```

#### set color vectors 
```{r set color vector}
coltimepoint <- c("#440154FF", "#3B528BFF", "#21908CFF", "#5DC863FF")
names(coltimepoint) <- c("E18", "P7", "3w", "8w")

collocation <- c("#61baba", "#ba6161")
names(collocation) <- c("iLN", "mLN")
```

### load object all
```{r}
basedir <- here()
fileNam <- paste0(basedir, "/data/LNmLToRev_allmerged_seurat.rds")
seuratM <- readRDS(fileNam)
table(seuratM$timepoint)
table(seuratM$orig.ident)
```

### subset E18
```{r}
seuratA <- subset(seuratM, timepoint == "E18")
table(seuratA$timepoint)

seuratA <- JoinLayers(seuratA)

#rerun seurat
seuratA <- NormalizeData (object = seuratA)
seuratA <- FindVariableFeatures(object = seuratA)
seuratA <- ScaleData(object = seuratA, verbose = TRUE)
seuratA <- RunPCA(object=seuratA, npcs = 30, verbose = FALSE)
seuratA <- RunTSNE(object=seuratA, reduction="pca", dims = 1:20)
seuratA <- RunUMAP(object=seuratA, reduction="pca", dims = 1:20)
seuratA <- FindNeighbors(object = seuratA, reduction = "pca", dims= 1:20)

res <- c(0.25, 0.6, 0.8, 0.4)
for (i in 1:length(res)) {
  seuratA <- FindClusters(object = seuratA, resolution = res[i], random.seed = 1234)
}
```

### dimplot all
#### clustering
```{r umap clustering pre filter, eval=FALSE, include=TRUE}
Idents(seuratA) <- seuratA$RNA_snn_res.0.25
DimPlot(seuratA, reduction = "umap", group.by = "RNA_snn_res.0.25" ,
        pt.size = 0.1, label = T, shuffle = T) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("umap1") +
  ylab("umap2")
```

#### location
```{r umap location pre filter, include=TRUE, eval=FALSE}
DimPlot(seuratA, reduction = "umap", group.by = "location" ,
        pt.size = 0.1, label = T, shuffle = T) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("umap1") +
  ylab("umap2")
```
### features
```{r featureplot sel genes pre filter}
genes <- data.frame(gene=rownames(seuratA)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) 

selGenes <- data.frame(geneID=c("Ptprc", "Msln", "Mki67", "Kcnn3", "Tcf21", "Pecam1", "Lyve1", "Ccl21a", "Icam1", "Cd34")) %>% 
  left_join(., genes, by = "geneID") 

pList <- sapply(selGenes$gene, function(x){
  p <- FeaturePlot(seuratA, reduction = "umap", 
            features = x, 
            cols=c("lightgrey","darkred"),
            order = T)+
  theme(legend.position="right")
  plot(p)
})
```

### filter 
```{r filter object, eval=FALSE, include=TRUE}
## filter Ptprc+ cells (cluster #7 and #14)
table(seuratA$RNA_snn_res.0.25)
seuratF <- subset(seuratA, RNA_snn_res.0.25 %in% c("7", "14"), invert = TRUE)
table(seuratF$RNA_snn_res.0.25)

seuratE18fil <- seuratF
remove(seuratF)
```

### rerun after fil
```{r rerun after fil, eval=FALSE, include=TRUE}
#rerun seurat
seuratE18fil <- NormalizeData (object = seuratE18fil)
seuratE18fil <- FindVariableFeatures(object = seuratE18fil)
seuratE18fil <- ScaleData(object = seuratE18fil, verbose = TRUE)
seuratE18fil <- RunPCA(object=seuratE18fil, npcs = 30, verbose = FALSE)
seuratE18fil <- RunTSNE(object=seuratE18fil, reduction="pca", dims = 1:20)
seuratE18fil <- RunUMAP(object=seuratE18fil, reduction="pca", dims = 1:20)
seuratE18fil <- FindNeighbors(object = seuratE18fil, reduction = "pca", dims= 1:20)

res <- c(0.25, 0.6, 0.8, 0.4)
for (i in 1:length(res)) {
  seuratE18fil <- FindClusters(object = seuratE18fil, resolution = res[i], random.seed = 1234)
}
```

### load object fil
```{r load object E18fil}
fileNam <- paste0(basedir, "/data/LNmLToRev_E18fil_seurat.rds")
seuratE18fil <- readRDS(fileNam)
```

### dimplot E18 fil
#### clustering
```{r dimplot clustering E18fil}
Idents(seuratE18fil) <- seuratE18fil$RNA_snn_res.0.25
colPal <- c("#DAF7A6", "#FFC300", "#FF5733", "#C70039", "#900C3F", "#b66e8d",
            "#61a4ba", "#6178ba", "#54a87f",  "#25328a",
            "#b6856e", "#0073C2FF", "#EFC000FF", "#868686FF", "#CD534CFF", 
            "#7AA6DCFF", "#003C67FF", "#8F7700FF", "#3B3B3BFF", "#A73030FF",
            "#4A6990FF")[1:length(unique(seuratE18fil$RNA_snn_res.0.25))]
names(colPal) <- unique(seuratE18fil$RNA_snn_res.0.25)

DimPlot(seuratE18fil, reduction = "umap", group.by = "RNA_snn_res.0.25",
        cols = colPal, label = TRUE)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")
```

#### location
```{r dimplot loction E18fil}
DimPlot(seuratE18fil, reduction = "umap", group.by = "location",
        cols = collocation)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")
```

```{r, eval=FALSE, eval=FALSE, include=TRUE}
saveRDS(seuratE18fil, file=paste0(basedir,"/data/LNmLToRev_E18fil_seurat.rds"))
```

### subset EYFP expressing cells
```{r subset EYFP pos cells, eval=FALSE, include=TRUE}
seuratSub <- subset(seuratE18fil, Rosa26eyfp.Rosa26eyfp>0)
eyfpPos <- colnames(seuratSub)

seuratE18fil$EYFP <- "neg"
seuratE18fil$EYFP[which(colnames(seuratE18fil)%in%eyfpPos)] <- "pos"
table(seuratE18fil$dataset, seuratE18fil$EYFP)

table(seuratE18fil$EYFP)
seuratE18EYFPv2 <- subset(seuratE18fil, EYFP == "pos")
table(seuratE18EYFPv2$EYFP)

DimPlot(seuratE18EYFPv2, reduction = "umap", group.by = "RNA_snn_res.0.25",
        cols = colPal, label = TRUE)

#rerun seurat
seuratE18EYFPv2 <- NormalizeData (object = seuratE18EYFPv2)
seuratE18EYFPv2<- FindVariableFeatures(object = seuratE18EYFPv2)
seuratE18EYFPv2 <- ScaleData(object = seuratE18EYFPv2, verbose = TRUE)
seuratE18EYFPv2 <- RunPCA(object=seuratE18EYFPv2, npcs = 30, verbose = FALSE)
seuratE18EYFPv2 <- RunTSNE(object=seuratE18EYFPv2, reduction="pca", dims = 1:20)
seuratE18EYFPv2 <- RunUMAP(object=seuratE18EYFPv2, reduction="pca", dims = 1:20)
seuratE18EYFPv2 <- FindNeighbors(object = seuratE18EYFPv2, reduction = "pca", dims= 1:20)

res <- c(0.25, 0.6, 0.8, 0.4)
for (i in 1:length(res)) {
  seuratE18EYFPv2 <- FindClusters(object = seuratE18EYFPv2, resolution = res[i], random.seed = 1234)
}
```

```{r, eval=FALSE, include=TRUE}
saveRDS(seuratE18EYFPv2, file=paste0(basedir,"/data/E18_EYFPv2_seurat.rds")
```

### load object E18 EYFP+
```{r}
fileNam <- paste0(basedir, "/data/E18_EYFPv2_seurat.rds")
seuratE18EYFPv2 <- readRDS(fileNam)
```

### dimplot E18 EYFP+
#### clustering
```{r dimplot clustering E18 EYFP pos v2}
Idents(seuratE18EYFPv2) <- seuratE18EYFPv2$RNA_snn_res.0.25
colPal <- c("#DAF7A6", "#FFC300", "#FF5733", "#C70039", "#900C3F", "#b66e8d",
            "#61a4ba", "#6178ba", "#54a87f",  "#25328a",
            "#b6856e", "#0073C2FF", "#EFC000FF", "#868686FF", "#CD534CFF", 
            "#7AA6DCFF", "#003C67FF", "#8F7700FF", "#3B3B3BFF", "#A73030FF",
            "#4A6990FF")[1:length(unique(seuratE18EYFPv2$RNA_snn_res.0.25))]
names(colPal) <- unique(seuratE18EYFPv2$RNA_snn_res.0.25)

DimPlot(seuratE18EYFPv2, reduction = "umap", group.by = "RNA_snn_res.0.25",
        cols = colPal, label = TRUE)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")
```

### location
```{r dimplot loction E18 EYFP v2}
DimPlot(seuratE18EYFPv2, reduction = "umap", group.by = "location",
        cols = collocation)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")
```

### features E18 EYFP+
```{r featureplot E18 EYFP only v2}
genes <- data.frame(gene=rownames(seuratE18EYFPv2)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) 

selGenes <- data.frame(geneID=c("Rosa26eyfp", "Mki67", "Acta2", "Myh11", "Ccl19", "Cxcl13", "Cd34", "Icam1","Vcam1", "Pi16")) %>% 
  left_join(., genes, by = "geneID") 

pList <- sapply(selGenes$gene, function(x){
  p <- FeaturePlot(seuratE18EYFPv2, reduction = "umap", 
            features = x, 
            cols=c("lightgrey","darkred"),
            order = T)+
  theme(legend.position="right")
  plot(p)
})
```

### integrate data across location 
```{r integrate v2, eval=FALSE, include=TRUE}
Idents(seuratE18EYFPv2) <- seuratE18EYFPv2$location

seurat.list <- SplitObject(object = seuratE18EYFPv2, split.by = "location")
for (i in 1:length(x = seurat.list)) {
    seurat.list[[i]] <- NormalizeData(object = seurat.list[[i]],
                                      verbose = FALSE)
    seurat.list[[i]] <- FindVariableFeatures(object = seurat.list[[i]], 
        selection.method = "vst", nfeatures = 2000, verbose = FALSE)
}

seurat.anchors <- FindIntegrationAnchors(object.list = seurat.list, dims = 1:20)
seuratE18EYFPv2.int <- IntegrateData(anchorset = seurat.anchors, dims = 1:20)
DefaultAssay(object = seuratE18EYFPv2.int) <- "integrated"

## rerun seurat
seuratE18EYFPv2.int <- ScaleData(object = seuratE18EYFPv2.int, verbose = FALSE,
                         features = rownames(seuratE18EYFPv2.int))
seuratE18EYFPv2.int <- RunPCA(object = seuratE18EYFPv2.int, npcs = 20, verbose = FALSE)
seuratE18EYFPv2.int <- RunTSNE(object = seuratE18EYFPv2.int, recuction = "pca", dims = 1:20)
seuratE18EYFPv2.int <- RunUMAP(object = seuratE18EYFPv2.int, recuction = "pca", dims = 1:20)

seuratE18EYFPv2.int <- FindNeighbors(object = seuratE18EYFPv2.int, reduction = "pca", dims = 1:20)
res <- c(0.6, 0.8, 0.4, 0.25)
for (i in 1:length(res)){
  seuratE18EYFPv2.int <- FindClusters(object = seuratE18EYFPv2.int, resolution = res[i],
                              random.seed = 1234)
}
```

### load object E18 EYFP+ integrated
```{r}
fileNam <- paste0(basedir, "/data/E18EYFPv2_integrated_seurat.rds")
seuratE18EYFPv2.int <- readRDS(fileNam)
```

```{r set default assay int object v2}
DefaultAssay(object = seuratE18EYFPv2.int) <- "RNA"
seuratE18EYFPv2.int$intCluster <- seuratE18EYFPv2.int$integrated_snn_res.0.25
Idents(seuratE18EYFPv2.int) <- seuratE18EYFPv2.int$intCluster

colPal <- c("#DAF7A6", "#FFC300", "#FF5733", "#C70039", "#900C3F", "#b66e8d",
            "#61a4ba", "#6178ba", "#54a87f", "#25328a", "#b6856e",
            "#ba6161", "#20714a", "#0073C2FF", "#EFC000FF", "#868686FF", 
            "#CD534CFF","#7AA6DCFF", "#003C67FF", "#8F7700FF", "#3B3B3BFF",
            "#A73030FF", "#4A6990FF")[1:length(unique(seuratE18EYFPv2.int$intCluster))]
names(colPal) <- unique(seuratE18EYFPv2.int$intCluster)
```

### dimplot E18 EYFP+ int
#### clustering 
```{r umap int v2}
DimPlot(seuratE18EYFPv2.int, reduction = "umap",
        label = T, shuffle = T, cols = colPal) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("umap1") +
  ylab("umap2")
```

#### location
```{r umap int location v2}
DimPlot(seuratE18EYFPv2.int, reduction = "umap", group.by = "location", cols = collocation,
        shuffle = T) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("umap1") +
  ylab("umap2")
```

### features E18 EYFP int
```{r featureplot E18 EYFP only int v2}
genes <- data.frame(gene=rownames(seuratE18EYFPv2.int)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) 

selGenes <- data.frame(geneID=c("Rosa26eyfp", "Mki67", "Acta2", "Myh11", "Mcam", "Ccl19", "Cxcl13", "Cd34", "Icam1","Vcam1", "Pi16", "Bmp4", "Fmod", "Adipoq", "Msln", "Kcnn3")) %>% 
  left_join(., genes, by = "geneID") 

pList <- sapply(selGenes$gene, function(x){
  p <- FeaturePlot(seuratE18EYFPv2.int, reduction = "umap", 
            features = x, 
            cols=c("lightgrey","darkred"),
            order = T)+
  theme(legend.position="right")
  plot(p)
})
```

#### assign label
```{r assign label int clusters v2}
seuratE18EYFPv2.int$label <- "label"
seuratE18EYFPv2.int$label[which(seuratE18EYFPv2.int$intCluster == "0")] <- "cluster2"
seuratE18EYFPv2.int$label[which(seuratE18EYFPv2.int$intCluster == "1")] <- "cluster3"
seuratE18EYFPv2.int$label[which(seuratE18EYFPv2.int$intCluster == "2")] <- "Prolif"
seuratE18EYFPv2.int$label[which(seuratE18EYFPv2.int$intCluster == "3")] <- "cluster1"
seuratE18EYFPv2.int$label[which(seuratE18EYFPv2.int$intCluster == "4")] <- "cluster4"
seuratE18EYFPv2.int$label[which(seuratE18EYFPv2.int$intCluster == "5")] <- "Neuronal1"
seuratE18EYFPv2.int$label[which(seuratE18EYFPv2.int$intCluster == "6")] <- "Mesothelial"
seuratE18EYFPv2.int$label[which(seuratE18EYFPv2.int$intCluster == "7")] <- "Neuronal2"
seuratE18EYFPv2.int$label[which(seuratE18EYFPv2.int$intCluster == "8")] <- "cluster5"
table(seuratE18EYFPv2.int$label)

##order
seuratE18EYFPv2.int$label <- factor(seuratE18EYFPv2.int$label, levels = c("cluster1", "cluster2", "cluster3", "cluster4", "cluster5", "Neuronal1","Neuronal2", "Mesothelial", "Prolif"))
table(seuratE18EYFPv2.int$label)

colLab <- c("#900C3F","#b66e8d", "#003C67FF",
            "#e3953d", "#714542", "#b6856e", "lightblue","grey", "black")
names(colLab) <- c("cluster1", "cluster2", "cluster3", "cluster4", "cluster5", "Neuronal1","Neuronal2", "Mesothelial", "Prolif")
```

#### label
```{r umap int label v2}
DimPlot(seuratE18EYFPv2.int, reduction = "umap", group.by = "label", cols = colLab)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seuratE18EYFPv2.int, reduction = "umap", group.by = "label", pt.size=0.5,
        cols = colLab, shuffle = T)+
  theme_void()

DimPlot(seuratE18EYFPv2.int, reduction = "umap", group.by = "label", pt.size=0.5,
        cols = colLab, shuffle = T)+
  theme_void() + 
  theme(legend.position = "none") 
```

### dotplot FRC marker E18 EYFP+ int
```{r dotplot FRC marker v2, fig.height=8, fig.width=6}
seurat_markers <- data.frame(gene=c("Fcgr2b","Fcer2a","Cr2","Cxcl13",
                                    "Slc7a11", "Ccl19", 
                                    "Ccl21a", "Fmod", "Grem1", "Bmp4",
                                    "Tnfsf11", "Fbn2",
                                    "Pltp" ,"C1rb", "Lepr", "Ptn",
                                     "Nr4a1", "Cxcl10", "Cxcl9", 
                                    "F3", "Fbln1", "Gdf10", "Adamtsl1", 
                                    "Col15a1",  "Cd34",
                                    "Igfbp6", "Pi16", "Thy1", "Dpp4", "Sema3c", 
                                    "Acta2", "Myh11", "Mcam", "Itga7", "Esam", "Rgs4", "Adipoq", "Mki67", "Msln", "Kcnn3", "Tcf21"
                                    ))

genes <- data.frame(geneID=rownames(seuratE18EYFPv2.int)) %>%
  mutate(gene=gsub(".*\\.", "",  geneID)) 

markerAll <- seurat_markers %>% left_join(., genes, by="gene")


## Dotplot all
Idents(seuratE18EYFPv2.int) <- seuratE18EYFPv2.int$label

DotPlot(seuratE18EYFPv2.int, assay="RNA", features = rev(markerAll$geneID), scale =T,
        cluster.idents = F) +
  scale_color_viridis_c() +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_discrete(breaks=rev(markerAll$geneID), labels=rev(markerAll$gene)) +
  xlab("") + ylab("")
```

```{r, eval=FALSE, include=TRUE}
saveRDS(seuratE18EYFPv2.int, file=paste0(basedir,"/data/E18EYFPv2_integrated_seurat.rds")
```

### subset FRCs and rerun
```{r}
table(seuratE18EYFPv2.int$label)
seuratE18EYFPv2.int <- subset(seuratE18EYFPv2.int, label %in% c("Neuronal1", "Neuronal2", "Mesothelial"), invert = TRUE)
table(seuratE18EYFPv2.int$label)

## rerun seurat
DefaultAssay(object = seuratE18EYFPv2.int) <- "integrated"
seuratE18EYFPv2.int <- ScaleData(object = seuratE18EYFPv2.int, verbose = FALSE,
                         features = rownames(seuratE18EYFPv2.int))
seuratE18EYFPv2.int <- RunPCA(object = seuratE18EYFPv2.int, npcs = 20, verbose = FALSE)
seuratE18EYFPv2.int <- RunTSNE(object = seuratE18EYFPv2.int, recuction = "pca", dims = 1:20)
seuratE18EYFPv2.int <- RunUMAP(object = seuratE18EYFPv2.int, recuction = "pca", dims = 1:20)

seuratE18EYFPv2.int <- FindNeighbors(object = seuratE18EYFPv2.int, reduction = "pca", dims = 1:20)
res <- c(0.1, 0.6, 0.8, 0.4, 0.25)
for (i in 1:length(res)){
  seuratE18EYFPv2.int <- FindClusters(object = seuratE18EYFPv2.int, resolution = res[i],
                              random.seed = 1234)
}
```

```{r set default assay int object v2 mLto only}
DefaultAssay(object = seuratE18EYFPv2.int) <- "RNA"
seuratE18EYFPv2.int$intCluster <- seuratE18EYFPv2.int$integrated_snn_res.0.1
Idents(seuratE18EYFPv2.int) <- seuratE18EYFPv2.int$intCluster

colPal <- c("#DAF7A6", "#FFC300", "#FF5733", "#C70039", "#900C3F", "#b66e8d",
            "#61a4ba", "#6178ba", "#54a87f", "#25328a", "#b6856e",
            "#ba6161", "#20714a", "#0073C2FF", "#EFC000FF", "#868686FF", 
            "#CD534CFF","#7AA6DCFF", "#003C67FF", "#8F7700FF", "#3B3B3BFF",
            "#A73030FF", "#4A6990FF")[1:length(unique(seuratE18EYFPv2.int$intCluster))]
names(colPal) <- unique(seuratE18EYFPv2.int$intCluster)
```

### dimplot E18 EYFP+ fil
#### clustering 
```{r umap int v2 mLto only}
DimPlot(seuratE18EYFPv2.int, reduction = "umap",
        label = T, shuffle = T, cols = colPal) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("umap1") +
  ylab("umap2")
```

#### location
```{r umap int location v2 mLTo only}
DimPlot(seuratE18EYFPv2.int, reduction = "umap", group.by = "location", cols = collocation,
        shuffle = T) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("umap1") +
  ylab("umap2")
```

#### assign label
```{r assign label int clusters v2 mLTo only}
seuratE18EYFPv2.int$label <- "label"
seuratE18EYFPv2.int$label[which(seuratE18EYFPv2.int$intCluster == "0")] <- "cluster3"
seuratE18EYFPv2.int$label[which(seuratE18EYFPv2.int$intCluster == "1")] <- "cluster1"
seuratE18EYFPv2.int$label[which(seuratE18EYFPv2.int$intCluster == "2")] <- "cluster2"
seuratE18EYFPv2.int$label[which(seuratE18EYFPv2.int$intCluster == "3")] <- "cluster4"
seuratE18EYFPv2.int$label[which(seuratE18EYFPv2.int$intCluster == "4")] <- "cluster5"
table(seuratE18EYFPv2.int$label)

##order
seuratE18EYFPv2.int$label <- factor(seuratE18EYFPv2.int$label, levels = c("cluster2", "cluster3", "cluster4", "cluster5", "cluster1"))
table(seuratE18EYFPv2.int$label)

colLab <- c("#900C3F","#b66e8d", "#003C67FF",
            "#e3953d", "#714542", "#b6856e")
names(colLab) <- c("cluster2", "cluster3", "cluster1", "cluster4", "cluster5")
```

#### label
```{r umap int label v2 mLTo only}
DimPlot(seuratE18EYFPv2.int, reduction = "umap", group.by = "label", cols = colLab)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seuratE18EYFPv2.int, reduction = "umap", group.by = "label", pt.size=0.5,
        cols = colLab, shuffle = T)+
  theme_void()

DimPlot(seuratE18EYFPv2.int, reduction = "umap", group.by = "label", pt.size=0.5,
        cols = colLab, shuffle = T)+
  theme_void() + 
  theme(legend.position = "none") 
```

### dotplot marker E18 EYFP+ fil
```{r dotplot FRC marker v2 mLTo only, fig.height=10, fig.width=6}
seurat_markers <- data.frame(gene=c("Vcam1", "Icam1",
                                    "Cxcl13", "Ccl19", "Ccl21a","Tnfsf11", "Grem1","Ifitm1","Cxcl1","Ifitm3","Ccl2","Nfkbia","Des",
                                    "Mfap5","Cdkn1c","Akap12","Anxa2","Lox","Gsn","Basp1","Fndc1","Sparc","Col1a1","Fbn2","Nr4a1","Fbln1","Cd34","Pi16",
                                    "Fbln5","Tm4sf1", "Ntrk3", "Fhl1", "Rgs7bp", "Adamts2", "Mpped2", "Ramp1", "Pdgfrl", "Eln", "Hspb2","Mgp", "Actg2","Acta2", "Myh11", "Mcam", "Mki67", "Ccna2", "Cdca8", "Prc1", "Aurkb"))

genes <- data.frame(geneID=rownames(seuratE18EYFPv2.int)) %>%
  mutate(gene=gsub(".*\\.", "",  geneID)) 

markerAll <- seurat_markers %>% left_join(., genes, by="gene")


## Dotplot all
Idents(seuratE18EYFPv2.int) <- seuratE18EYFPv2.int$label

DotPlot(seuratE18EYFPv2.int, assay="RNA", features = rev(markerAll$geneID), scale =T,
        cluster.idents = F) +
  scale_color_viridis_c() +
  coord_flip() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_discrete(breaks=rev(markerAll$geneID), labels=rev(markerAll$gene)) +
  xlab("") + ylab("")
```

###signatures
#### convert to sce 
```{r convert to sce mLTo only}
## convert seurat object to sce object
## exteract logcounts
logcounts <- GetAssayData(seuratE18EYFPv2.int, assay = "RNA", slot = "data")
counts <- GetAssayData(seuratE18EYFPv2.int, assay = "RNA", slot = "counts")
## extract reduced dims from integrated assay
pca <- Embeddings(seuratE18EYFPv2.int, reduction = "pca")
umap <- Embeddings(seuratE18EYFPv2.int, reduction = "umap")
## create sce object
sce <- SingleCellExperiment(assays =list (
  counts = counts,
  logcounts = logcounts
),
colData = seuratE18EYFPv2.int@meta.data,
rowData = data.frame(gene_id = rownames(logcounts)),
reducedDims = SimpleList(
  PCA = pca,
  UMAP = umap
))

genes <- data.frame(geneID=rownames(sce)) %>% mutate(gene=gsub(".*\\.", "", geneID))
pal = colorRampPalette(c("#053061", "#2166ac", "#f7f7f7", "#f4a582", "#b2183c", "#85122d"))
```

### signatures
```{r}
selGenes <- data.frame(gene=c("Cxcl13", "Ccl19", "Ccl21a","Tnfsf11", "Grem1"))
signGenes <- genes %>% dplyr::filter(gene %in% selGenes$gene)

##make a count matrix of signature genes
sceSub <- sce[which(rownames(sce) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceSub@assays@data$logcounts)))/nrow(signGenes)
sceSub$sign <- cntMat
sceSub$sign2 <- sceSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 3))
sceSub$sign2[which(sceSub$sign > 3)] <- 3
##check max and min values
max(sceSub$sign)
min(sceSub$sign)
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc
```

```{r}
selGenes <- data.frame(gene=c("Mfap5","Gsn","Fndc1","Col1a1","Cd34"))
signGenes <- genes %>% dplyr::filter(gene %in% selGenes$gene)

##make a count matrix of signature genes
sceSub <- sce[which(rownames(sce) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceSub@assays@data$logcounts)))/nrow(signGenes)
sceSub$sign <- cntMat
sceSub$sign2 <- sceSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 3))
sceSub$sign2[which(sceSub$sign > 3)] <- 3
##check max and min values
max(sceSub$sign)
min(sceSub$sign)
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc
```

```{r}
selGenes <- data.frame(gene=c("Fbln5","Eln","Actg2","Acta2","Myh11"))
signGenes <- genes %>% dplyr::filter(gene %in% selGenes$gene)

##make a count matrix of signature genes
sceSub <- sce[which(rownames(sce) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceSub@assays@data$logcounts)))/nrow(signGenes)
sceSub$sign <- cntMat
sceSub$sign2 <- sceSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 3))
sceSub$sign2[which(sceSub$sign > 3)] <- 3
##check max and min values
max(sceSub$sign)
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc
```

### plot signature 3/4 combined
```{r}
selGenes <- data.frame(gene=c("Fbln5","Eln","Actg2","Acta2","Myh11","Mfap5","Gsn","Fndc1","Col1a1","Cd34"))
signGenes <- genes %>% dplyr::filter(gene %in% selGenes$gene)

##make a count matrix of signature genes
sceSub <- sce[which(rownames(sce) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceSub@assays@data$logcounts)))/nrow(signGenes)
sceSub$sign <- cntMat
sceSub$sign2 <- sceSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 2.5))
sceSub$sign2[which(sceSub$sign > 2.5)] <- 2.5
##check max and min values
max(sceSub$sign)
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc
```

```{r}
selGenes <- data.frame(gene=c("Mki67", "Ccna2", "Cdca8", "Prc1", "Aurkb"))
signGenes <- genes %>% dplyr::filter(gene %in% selGenes$gene)

##make a count matrix of signature genes
sceSub <- sce[which(rownames(sce) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceSub@assays@data$logcounts)))/nrow(signGenes)
sceSub$sign <- cntMat
sceSub$sign2 <- sceSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 2))
sceSub$sign2[which(sceSub$sign > 2)] <- 2
##check max and min values
max(sceSub$sign)
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc
```

#### featrue Ccl19
```{r feature Ccl19}
genes <- data.frame(gene=rownames(seuratE18EYFPv2.int)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) 

selGenes <- data.frame(geneID=c("Ccl19")) %>% 
  left_join(., genes, by = "geneID") 

pList <- sapply(selGenes$gene, function(x){
  p <- FeaturePlot(seuratE18EYFPv2.int, reduction = "umap", 
            features = x, 
            cols=c("lightgrey","darkred"),
            order = FALSE)+
  theme(legend.position="right")
  plot(p)
})
```


### subset Ccl19 positive cells
```{r}
seuratCcl19 <- subset(seuratE18EYFPv2.int, ENSMUSG00000071005.Ccl19 > 0)
table(seuratE18EYFPv2.int$orig.ident)
table(seuratCcl19$orig.ident)
```

#### rerun Ccl19 positive only
```{r}
## rerun seurat
DefaultAssay(object = seuratCcl19) <- "integrated"
seuratCcl19 <- ScaleData(object = seuratCcl19, verbose = FALSE,
                         features = rownames(seuratCcl19))
seuratCcl19 <- RunPCA(object = seuratCcl19, npcs = 20, verbose = FALSE)
seuratCcl19 <- RunTSNE(object = seuratCcl19, recuction = "pca", dims = 1:20)
seuratCcl19 <- RunUMAP(object = seuratCcl19, recuction = "pca", dims = 1:20)

seuratCcl19 <- FindNeighbors(object = seuratCcl19, reduction = "pca", dims = 1:20)
res <- c(0.1, 0.6, 0.8, 0.4, 0.25)
for (i in 1:length(res)){
  seuratCcl19 <- FindClusters(object = seuratCcl19, resolution = res[i],
                              random.seed = 1234)
}
```

```{r set default assay Ccl19}
DefaultAssay(object = seuratCcl19) <- "RNA"
Idents(seuratCcl19) <- seuratCcl19$integrated_snn_res.0.25

colPal <- c("#DAF7A6", "#FFC300", "#FF5733", "#C70039", "#900C3F", "#b66e8d",
            "#61a4ba", "#6178ba", "#54a87f", "#25328a", "#b6856e",
            "#ba6161", "#20714a", "#0073C2FF", "#EFC000FF", "#868686FF", 
            "#CD534CFF","#7AA6DCFF", "#003C67FF", "#8F7700FF", "#3B3B3BFF",
            "#A73030FF", "#4A6990FF")[1:length(unique(seuratCcl19$integrated_snn_res.0.25))]
names(colPal) <- unique(seuratCcl19$integrated_snn_res.0.25)
```

### dimplots Ccl19+ cells
#### clustering 
```{r}
DimPlot(seuratCcl19, reduction = "umap",
        label = T, shuffle = T, cols = colPal) +
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("umap1") +
  ylab("umap2")
```

#### assign label Ccl19+ cells
```{r assign label Ccl19}
seuratCcl19$label <- "label"
seuratCcl19$label[which(seuratCcl19$integrated_snn_res.0.25 == "0")] <- "cluster2"
seuratCcl19$label[which(seuratCcl19$integrated_snn_res.0.25 == "1")] <- "cluster1"
seuratCcl19$label[which(seuratCcl19$integrated_snn_res.0.25 == "2")] <- "cluster3"
table(seuratCcl19$label)

##order
seuratCcl19$label <- factor(seuratCcl19$label, levels = c("cluster2", "cluster3","cluster1"))
table(seuratCcl19$label)

colLab <- c("#900C3F","#b66e8d", "#003C67FF")
names(colLab) <- c("cluster2", "cluster3", "cluster1")
```



#### label
```{r}
DimPlot(seuratCcl19, reduction = "umap", group.by = "label", cols = colLab)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")

DimPlot(seuratCcl19, reduction = "umap", group.by = "label", pt.size=0.5,
        cols = colLab, shuffle = T)+
  theme_void()

DimPlot(seuratCcl19, reduction = "umap", group.by = "label", pt.size=0.5,
        cols = colLab, shuffle = T)+
  theme_void() + 
  theme(legend.position = "none") 
```

#### featrue plot Ccl19
```{r featureplot Ccl19+}
genes <- data.frame(gene=rownames(seuratCcl19)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) 

selGenes <- data.frame(geneID=c("Ccl19")) %>% 
  left_join(., genes, by = "geneID") 

pList <- sapply(selGenes$gene, function(x){
  p <- FeaturePlot(seuratCcl19, reduction = "umap", 
            features = x, 
            cols=c("lightgrey","darkred"),
            order = FALSE)+
  theme(legend.position="right")
  plot(p)
})
```

#### signatures Ccl19+ cells
```{r}
##convert seurat object to sce object
##exteract logcounts
logcounts <- GetAssayData(seuratCcl19, assay = "RNA", slot = "data")
counts <- GetAssayData(seuratCcl19, assay = "RNA", slot = "counts")
##extract reduced dims from integrated assay
pca <- Embeddings(seuratCcl19, reduction = "pca")
umap <- Embeddings(seuratCcl19, reduction = "umap")
##create sce object
sce <- SingleCellExperiment(assays =list (
  counts = counts,
  logcounts = logcounts
),
colData = seuratCcl19@meta.data,
rowData = data.frame(gene_id = rownames(logcounts)),
reducedDims = SimpleList(
  PCA = pca,
  UMAP = umap
))

genes <- data.frame(geneID=rownames(sce)) %>% mutate(gene=gsub(".*\\.", "", geneID))
pal = colorRampPalette(c("#053061", "#2166ac", "#f7f7f7", "#f4a582", "#b2183c", "#85122d"))
```

```{r}
selGenes <- data.frame(gene=c("Cxcl13","Ccl19", "Ccl21a","Tnfsf11", "Grem1"))
signGenes <- genes %>% dplyr::filter(gene %in% selGenes$gene)

##make a count matrix of signature genes
sceSub <- sce[which(rownames(sce) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceSub@assays@data$logcounts)))/nrow(signGenes)
sceSub$sign <- cntMat
sceSub$sign2 <- sceSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 2.5))
sceSub$sign2[which(sceSub$sign > 2.5)] <- 2.5
sceSub$sign2[which(sceSub$sign < 0)] <- 0
##check max and min values
max(sceSub$sign)
min(sceSub$sign)
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc
```

```{r}
selGenes <- data.frame(gene=c("Fbln5","Eln","Actg2","Acta2","Myh11","Mfap5","Gsn","Fndc1","Col1a1","Cd34"))
signGenes <- genes %>% dplyr::filter(gene %in% selGenes$gene)

##make a count matrix of signature genes
sceSub <- sce[which(rownames(sce) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceSub@assays@data$logcounts)))/nrow(signGenes)
sceSub$sign <- cntMat
sceSub$sign2 <- sceSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 2.5))
sceSub$sign2[which(sceSub$sign > 2.5)] <- 2.5
sceSub$sign2[which(sceSub$sign < 0)] <- 0
##check max and min values
max(sceSub$sign)
min(sceSub$sign)
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc
```

```{r}
selGenes <- data.frame(gene=c("Mki67", "Ccna2", "Cdca8", "Prc1", "Aurkb"))
signGenes <- genes %>% dplyr::filter(gene %in% selGenes$gene)

##make a count matrix of signature genes
sceSub <- sce[which(rownames(sce) %in% signGenes$geneID),]
cntMat <- rowSums(t(as.matrix(
    sceSub@assays@data$logcounts)))/nrow(signGenes)
sceSub$sign <- cntMat
sceSub$sign2 <- sceSub$sign
sc <- scale_colour_gradientn(colours = pal(100), limits=c(0, 2))
sceSub$sign2[which(sceSub$sign > 2)] <- 2
sceSub$sign2[which(sceSub$sign < 0)] <- 0
##check max and min values
max(sceSub$sign)
min(sceSub$sign)
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc +
  theme(legend.position = "none")
plotUMAP(sceSub, colour_by = "sign2", point_size = 1) + sc
```

## session info
```{r date and session info}
date()
sessionInfo()
```
