---
title: "mLN EYFP+ all timepoints"
author: "A.DeMartin"
date: "2025-04-29"
output: 
  html_document:
    keep_md: true
    toc: true
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
options(width = 100)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
seed <- 1234
```

#### load packages
```{r load packages, warning=FALSE}
library(ExploreSCdataSeurat3)
library(runSeurat3)
library(Seurat)
library(ggpubr)
library(pheatmap)
library(SingleCellExperiment)
library(dplyr)
library(tidyverse)
library(viridis)
library(muscat)
library(circlize)
library(destiny)
library(scater)
library(metap)
library(multtest)
library(clusterProfiler)
library(org.Hs.eg.db)
library(msigdbr)
library(enrichplot)
library(DOSE)
library(grid)
library(gridExtra)
library(ggupset)
library(VennDiagram)
library(NCmisc)
library(slingshot)
library(RColorBrewer)
library(here)
```

### preprocessing
#### load object all
```{r}
basedir <- here()
fileNam <- paste0(basedir, "/data/LNmLToRev_allmerged_seurat.rds")
seuratM <- readRDS(fileNam)
table(seuratM$dataset)
table(seuratM$RNA_snn_res.0.25)
table(seuratM$orig.ident)
```

#### subset EYFP+ 
```{r subset EYFPpos, eval=FALSE, include=TRUE}
seuratSub <- subset(seuratM, Rosa26eyfp.Rosa26eyfp>0)
eyfpPos <- colnames(seuratSub)

seuratM$EYFP <-"neg"
seuratM$EYFP[which(colnames(seuratM)%in%eyfpPos)] <- "pos"

table(seuratM$dataset, seuratM$EYFP)

seuratEYFP <- subset(seuratM, EYFP == "pos")
table(seuratEYFP$orig.ident)

## rerun seurat
seuratEYFP <- NormalizeData (object = seuratEYFP)
seuratEYFP <- FindVariableFeatures(object = seuratEYFP)
seuratEYFP <- ScaleData(object = seuratEYFP, verbose = TRUE)
seuratEYFP <- RunPCA(object=seuratEYFP, npcs = 30, verbose = FALSE)
seuratEYFP <- RunTSNE(object=seuratEYFP, reduction="pca", dims = 1:20)
seuratEYFP <- RunUMAP(object=seuratEYFP, reduction="pca", dims = 1:20)
seuratEYFP <- FindNeighbors(object = seuratEYFP, reduction = "pca", dims= 1:20)

res <- c(0.25, 0.6, 0.8, 0.4)
for (i in 1:length(res)) {
  seuratEYFP <- FindClusters(object = seuratEYFP, resolution = res[i], random.seed = 1234)
}
```

```{r, eval=FALSE, include=TRUE}
## save object
saveRDS(seuratEYFP, file=paste0(basedir,"/data/LNmLToRev_EYFPonly_seurat.rds"))
```

### load object EYFP+
```{r}
fileNam <- paste0(basedir, "/data/LNmLToRev_EYFPonly_seurat.rds")
seuratEYFP <- readRDS(fileNam)
table(seuratEYFP$dataset)
table(seuratEYFP$RNA_snn_res.0.25)
table(seuratEYFP$orig.ident)
```

#### subset mLN
```{r, eval=FALSE, include=TRUE}
## subset
table(seuratEYFP$location)
seuratmLN <- subset(seuratEYFP, location == "mLN")
table(seuratmLN$orig.ident)

## rerun seurat
seuratmLN <- NormalizeData (object = seuratmLN)
seuratmLN <- FindVariableFeatures(object = seuratmLN)
seuratmLN <- ScaleData(object = seuratmLN, verbose = TRUE)
seuratmLN <- RunPCA(object=seuratmLN, npcs = 30, verbose = FALSE)
seuratmLN <- RunTSNE(object=seuratmLN, reduction="pca", dims = 1:20)
seuratmLN <- RunUMAP(object=seuratmLN, reduction="pca", dims = 1:20)
seuratmLN <- FindNeighbors(object = seuratmLN, reduction = "pca", dims= 1:20)

res <- c(0.25, 0.6, 0.8, 0.4)
for (i in 1:length(res)) {
  seuratmLN <- FindClusters(object = seuratmLN, resolution = res[i], random.seed = 1234)
}
```

```{r, eval=FALSE, include=TRUE}
### save object
saveRDS(seuratmLN, file=paste0(basedir,"/data/LNmLToRev_EYFP_mLN_seurat.rds"))
```

### load object mLN EYFP+
```{r}
fileNam <- paste0(basedir, "/data/LNmLToRev_EYFP_mLN_seurat.rds")
seuratmLN <- readRDS(fileNam)
table(seuratmLN$dataset)
table(seuratmLN$RNA_snn_res.0.25)
table(seuratmLN$orig.ident)
```

#### set color vectors 
```{r set color vector}
coltimepoint <- c("#440154FF", "#3B528BFF", "#21908CFF", "#5DC863FF")
names(coltimepoint) <- c("E18", "P7", "3w", "8w")

colPal <- c("#DAF7A6", "#FFC300", "#FF5733", "#C70039", "#900C3F", "#b66e8d",
            "#61a4ba", "#6178ba", "#54a87f",  "#25328a",
            "#b6856e", "#0073C2FF", "#EFC000FF", "#868686FF", "#CD534CFF", 
            "#7AA6DCFF", "#003C67FF", "#8F7700FF", "#3B3B3BFF", "#A73030FF",
            "#4A6990FF")[1:length(unique(seuratmLN$RNA_snn_res.0.4))]
names(colPal) <- unique(seuratmLN$RNA_snn_res.0.4)

collocation <- c("#61baba", "#ba6161")
names(collocation) <- c("iLN", "mLN")
```

### dimplot
#### clustering
```{r dimplot clustering}
DimPlot(seuratmLN, reduction = "umap", group.by = "RNA_snn_res.0.4",
        cols = colPal, label = TRUE)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")
```

#### timepoint
```{r dimplot age}
DimPlot(seuratmLN, reduction = "umap", group.by = "timepoint",
        cols = coltimepoint)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")
```

#### calculate cluster marker genes
```{r marker RNA_snn_res.0.4, include=TRUE, eval=FALSE}
## cluster marker
Idents(seuratmLN) <- seuratmLN$RNA_snn_res.0.4
markerGenes <- FindAllMarkers(seuratmLN, only.pos=T) %>% 
  dplyr::filter(p_val_adj < 0.01)
```

#### features
```{r featureplot sel genes}
genes <- data.frame(gene=rownames(seuratmLN)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) 

selGenesAll <- data.frame(geneID=c("Clu", "Ccl21a", "Grem1", "Grem2","Mfap4", "Ccl19","Cd34", "Col6a1",
                    "Col6a2","Rosa26eyfp", "Cxcl13", "Icam1", "Vcam1")) %>% 
  left_join(., genes, by = "geneID") 

pList <- sapply(selGenesAll$gene, function(x){
p <- FeaturePlot(seuratmLN, reduction = "umap", 
            features = x,
            cols=c("lightgrey", "darkred"),
            order = F)+
  theme(legend.position="right")
  plot(p)
})
```

```{r featureplot sel genes-2}
selGenesAll <- data.frame(geneID=c("Fbln1", "Col15a1", "Cnn1", "Acta2", "Rgs5",
                                   "Cox4i2", "Pi16", "Cd34", "Emp1", "Ogn",
                                   "Fhl2")) %>% 
  left_join(., genes, by = "geneID") 

pList <- sapply(selGenesAll$gene, function(x){
p <- FeaturePlot(seuratmLN, reduction = "umap", 
            features = x,
            cols=c("lightgrey", "darkred"),
            order = F)+
  theme(legend.position="right")
  plot(p)
})
```

```{r}
selGenesfil <- c("ENSMUSG00000026395.Ptprc", "ENSMUSG00000031004.Mki67", "ENSMUSG00000063011.Msln", "ENSMUSG00000045680.Tcf21")

pList <- sapply(selGenesfil, function(x){
  p <- FeaturePlot(seuratmLN, reduction = "umap", 
            features = x, 
            cols=c("lightgrey", "darkred"),
            order = T)+
  theme(legend.position="right")
  plot(p)
})
```

### filter 
```{r filter object 3, eval=FALSE, include=TRUE}
## filter out Ptprc+ cells (cluster #12), 
## and pancreatic cells (#15),
## and mesothelial cells (cluster#8) 
## and epithelial/neuronal cells (#11)
## and proliferating cells (cluster #5 and #7)

table(seuratmLN$RNA_snn_res.0.4)
seuratF <- subset(seuratmLN, RNA_snn_res.0.4 %in% c("12", "15", "8", "11", "5", "7"), invert = TRUE)
table(seuratF$RNA_snn_res.0.4)

seuratmLNf3 <- seuratF
remove(seuratF)
table(seuratmLNf3$orig.ident)
```

#### rerun 
```{r rerun mLNf3, eval=FALSE, include=TRUE}
## rerun seurat
seuratmLNf3 <- NormalizeData (object = seuratmLNf3)
seuratmLNf3 <- FindVariableFeatures(object = seuratmLNf3)
seuratmLNf3 <- ScaleData(object = seuratmLNf3, verbose = TRUE)
seuratmLNf3 <- RunPCA(object=seuratmLNf3, npcs = 30, verbose = FALSE)
seuratmLNf3 <- RunTSNE(object=seuratmLNf3, reduction="pca", dims = 1:20)
seuratmLNf3 <- RunUMAP(object=seuratmLNf3, reduction="pca", dims = 1:20)
seuratmLNf3 <- FindNeighbors(object = seuratmLNf3, reduction = "pca", dims= 1:20)

res <- c(0.25, 0.6, 0.8, 0.4)
for (i in 1:length(res)) {
  seuratmLNf3 <- FindClusters(object = seuratmLNf3, resolution = res[i], random.seed = 1234)
}
```

### load object mLN EYFP+ fil
```{r load object mLNf3}
fileNam <- paste0(basedir, "/data/LNmLToRev_EYFP_mLNf3_seurat.rds")
seuratmLNf3 <- readRDS(fileNam)
table(seuratmLNf3$dataset)
```

### dimplot
#### clustering
```{r dimplot clustering mLNf3}
colPal <- c("#DAF7A6", "#FFC300", "#FF5733", "#C70039", "#900C3F", "#b66e8d",
            "#61a4ba", "#6178ba", "#54a87f",  "#25328a",
            "#b6856e", "#0073C2FF", "#EFC000FF", "#868686FF", "#CD534CFF", 
            "#7AA6DCFF", "#003C67FF", "#8F7700FF", "#3B3B3BFF", "#A73030FF",
            "#4A6990FF")[1:length(unique(seuratmLNf3$RNA_snn_res.0.4))]
names(colPal) <- unique(seuratmLNf3$RNA_snn_res.0.4)

DimPlot(seuratmLNf3, reduction = "umap", group.by = "RNA_snn_res.0.4",
        cols = colPal, label = TRUE)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")
```

#### timepoint
```{r dimplot timepoint mLNf3}
DimPlot(seuratmLNf3, reduction = "umap", group.by = "timepoint",
        cols = coltimepoint)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")
```

### label transfer
```{r transfer label}
fileNam <- paste0(basedir, "/data/LNmLToRev_adultonly_seurat.integrated.rds")
seuratLab <- readRDS(fileNam)

seuratLab <- subset(seuratLab, location=="mLN")
seuratLab <- subset(seuratLab, EYFP=="pos")
table(seuratLab$label)
```

#### dimplot label
```{r transfer label-2}
labCells <- data.frame(label=seuratLab$label) %>% rownames_to_column(., "cell")
allCell <- data.frame(cell=colnames(seuratmLNf3)) %>% 
  left_join(., labCells, by= "cell")
allCell$label[which(is.na(allCell$label))] <- "unassigned"
seuratmLNf3$label <- allCell$label

table(seuratmLNf3$timepoint)
table(seuratmLNf3$label)

colLab <- c("#42a071", "#900C3F","#b66e8d", "#61a4ba", "#424671", "#003C67FF",
            "#e3953d", "#714542", "#b6856e", "#a4a4a4")

names(colLab) <- c("FDC/MRC", "TRC", "TBRC", "MedRC/IFRC", "MedRC" , "actMedRC",
                   "PRC", "Pi16+RC", "VSMC", "unassigned")

DimPlot(seuratmLNf3, reduction = "umap", group.by = "label",
        cols = colLab, shuffle=T)+
  theme_bw() +
  theme(axis.text = element_blank(), axis.ticks = element_blank(), 
        panel.grid.minor = element_blank()) +
  xlab("UMAP1") +
  ylab("UMAP2")
```

```{r transfer label-3}
DimPlot(seuratmLNf3, reduction = "umap", group.by = "label", pt.size=0.5,
        cols = colLab, order = TRUE)+
  theme_void()

DimPlot(seuratmLNf3, reduction = "umap", group.by = "label", pt.size=0.5,
        cols = colLab, shuffle = FALSE)+
  theme_void()
```

#### calculate cluster marker genes mLNf3
```{r marker RNA_snn_res.0.4 mLNf3, include=TRUE, eval=FALSE}
##cluster marker
Idents(seuratmLNf3) <- seuratmLNf3$RNA_snn_res.0.4
markerGenes <- FindAllMarkers(seuratmLNf3, only.pos=T) %>% 
  dplyr::filter(p_val_adj < 0.01)
```

### features
```{r vis markers mLNf3}
genes <- data.frame(gene=rownames(seuratmLNf3)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) 

selGenesAll <- data.frame(geneID=c("Rosa26eyfp","tdTomato", "Ccl19", "Ccl21a", "Cxcl13",
                                   "Fbln1", "Col15a1", "Cnn1", "Acta2","Myh11", "Rgs5",
                                   "Cox4i2", "Pi16", "Cd34", "Emp1", "Ogn","Des",
                                   "Fhl2", "Bmp2", "Bmp4", "Grem1", "Grem2", "Bmpr1a", "Bmpr1b", "Bmpr2")) %>% 
  left_join(., genes, by = "geneID") 

pList <- sapply(selGenesAll$gene, function(x){
p <- FeaturePlot(seuratmLNf3, reduction = "umap", 
            features = x,
            cols=c("lightgrey", "darkred"),
            order = F) +
  theme(legend.position="right")
  plot(p)
})
```

```{r save object mLNf3, eval=FALSE, include=TRUE}
## save object
saveRDS(seuratmLNf3, file=paste0(basedir,"/data/LNmLToRev_EYFP_mLNf3_seurat.rds"))
```

### slingshot
```{r slingshot mLNf3 v2, eval=FALSE, include=TRUE}
sce <- as.SingleCellExperiment(seuratmLNf3)
sce <- slingshot(sce, clusterLabels = 'RNA_snn_res.0.4', reducedDim = 'UMAP',
                 start.clus = "1", end.clus = c("12", "3", "4", "7", "6"),
                 dist.method="simple", extend = 'n', stretch=0)
```

#### load slingshot sce object mLN EYFP+
```{r load sce mLNf3 slingshot v2}
fileNam <- paste0(basedir, "/data/LNmLToRev_EYFP_mLNf3_slingshot_v2_sce.rds")
scemLNf3v2 <- readRDS(fileNam)
```

```{r slingshot mLNf3-2 v2}
clustDat <- data.frame(clustCol=colPal) %>% rownames_to_column(., "cluster")
timepointDat <- data.frame(ageCol=coltimepoint) %>% rownames_to_column(., "timepoint")
colDat <- data.frame(cluster=scemLNf3v2$RNA_snn_res.0.4) %>%
  mutate(timepoint=scemLNf3v2$timepoint) %>% left_join(., clustDat, by="cluster") %>% 
  left_join(., timepointDat, by="timepoint")
```

```{r slingshot-1 mLNf3 v2}
plot(reducedDims(scemLNf3v2)$UMAP, col = colDat$clustCol, pch=16, asp = 1)
lines(SlingshotDataSet(scemLNf3v2), lwd=2, type = 'lineages', col = 'black')
```

```{r slingshot-2 mLNf3 v2}
plot(reducedDims(scemLNf3v2)$UMAP, col = colDat$ageCol, pch=16, asp = 1)
lines(SlingshotDataSet(scemLNf3v2), lwd=2, type = 'lineages', col = 'black')
```

```{r slingshot-3 mLNf3 v2}
plot(reducedDims(scemLNf3v2)$UMAP, col = colDat$clustCol, pch=16, asp = 1)
lines(SlingshotDataSet(scemLNf3v2), lwd=2, col='black')
```

```{r slingshot-4 mLNf3 v2}
plot(reducedDims(scemLNf3v2)$UMAP, col = colDat$ageCol, pch=16, asp = 1)
lines(SlingshotDataSet(scemLNf3v2), lwd=2, col='black')
```

```{r save slingshot mLNf3 v2, eval=FALSE, include=TRUE}
saveRDS(sce, file=paste0(basedir,"/data/LNmLToRev_EYFP_mLNf3_slingshot_v2_sce.rds"))
```

```{r}
summary(scemLNf3v2$slingPseudotime_1)
summary(scemLNf3v2$slingPseudotime_2)
summary(scemLNf3v2$slingPseudotime_3)
summary(scemLNf3v2$slingPseudotime_4)
summary(scemLNf3v2$slingPseudotime_5)
```

```{r}
colors <- colorRampPalette(rev(brewer.pal(11,'Spectral')))(100)
plotcol <- colors[cut(slingAvgPseudotime(SlingshotDataSet(scemLNf3v2)), breaks=100)]

plot(reducedDims(scemLNf3v2)$UMAP, col = plotcol, pch=16, asp = 1)
lines(SlingshotDataSet(scemLNf3v2), lwd=2, col='black')
```

```{r}
colors <- colorRampPalette(brewer.pal(11,'YlOrRd'))(100)
plotcol <- colors[cut(slingAvgPseudotime(SlingshotDataSet(scemLNf3v2)), breaks=100)]

plot(reducedDims(scemLNf3v2)$UMAP, col = plotcol, pch=16, asp = 1)
lines(SlingshotDataSet(scemLNf3v2), lwd=2, col='black')
```

```{r}
colors <- colorRampPalette(brewer.pal(11,'YlGnBu'))(100)
plotcol <- colors[cut(slingAvgPseudotime(SlingshotDataSet(scemLNf3v2)), breaks=100)]

plot(reducedDims(scemLNf3v2)$UMAP, col = plotcol, pch=16, asp = 1)
lines(SlingshotDataSet(scemLNf3v2), lwd=2, col='black')
```

```{r}
colors <- colorRampPalette(brewer.pal(11,'PuOr')[-6])(100)
plotcol <- colors[cut(slingAvgPseudotime(SlingshotDataSet(scemLNf3v2)), breaks=100)]

plot(reducedDims(scemLNf3v2)$UMAP, col = plotcol, pch=16, asp = 1)
lines(SlingshotDataSet(scemLNf3v2), lwd=2, col='black')
```


```{r}
### color lineages
colLin <- c("#42a071","#900C3F","#424671","#e3953d","#b6856e")
names(colLin) <- c("1", "2", "3", "4", "5")

plot(reducedDims(scemLNf3v2)$UMAP, col = "#bfbcbd", pch=16, asp = 1)
lines(SlingshotDataSet(scemLNf3v2), lwd=4, col=colLin)
```

### session info
```{r date and session info}
date()
sessionInfo()
```
